---
title: "Predicting Patriot League Standings Based on Conference Games (Updated After Every Game)"
author: "Noah Seaner"
date: "2/23/2026"
output: html_document
---
# Introduction

## The Data

The data we will be looking at in this updated simulation comes from every conference game that a Patriot League team has played so far this season ($16$ for each of the $10$ teams). The specific metrics are the same - for each game so far, I recorded the amount of points a team scored and allowed, as well as their field goal attempts, three-point attempts, free throw attempts, offensive rebounds, and turnovers in each game. Then, I estimated possessions, offensive rating, and defensive rating using the same formulas as the other simulation. Lastly, I calculated the means and standard deviations of these three measures, which we will use to simulate the remaining matchups on the schedule.

As the Patriot League plays more games, these numbers will become more statistically sound, whereas at the start, we only had five games worth of data from each team. I'll be doing updates after every game from now until the end of the season, with more statistically based numbers likely resulting around the halfway point of the conference schedule (around game $9$).

## Methodology

As was the case in the previous simulation, we will be sticking with the Monte Carlo method to simulate head-to-head matchups as well as the remaining portion of the schedule. We'll be running $10,000$ simulations of each team's remaining conference games to see how many wins they will accrue, as well as the head-to-head win probabilities for the remaining games.

The simulation code is the exact same as in the non-conference simulation, so I will only be showing the results of each team's simulation here for simplicity.

# The Simulation

## American

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "American") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `American Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `American Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Army

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Army") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Army Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Army Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Boston

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Boston") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Boston Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Boston Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Bucknell

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Bucknell") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Bucknell Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Bucknell Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Colgate

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Colgate") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Colgate Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Colgate Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Holy Cross

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Holy Cross") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Holy Cross Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Holy Cross Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Lafayette

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Lafayette") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Lafayette Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Lafayette Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Lehigh

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Lehigh") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Lehigh Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Lehigh Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Loyola

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Loyola") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Loyola Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Loyola Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

## Navy

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(stringr)

set.seed(123) 
n_sims <- 10000
home_court_adv <- 2.5 
min_possessions <- 55
max_possessions <- 85

games_hist <- read_excel("conf record sim update.xlsx", sheet = "PL Games") %>% 
  mutate(Date = as.Date(Date))

sched <- read_excel("conf record sim update.xlsx", sheet = "Navy") %>% 
  mutate(Date = as.Date(Date))

ratings <- read_excel("conf record sim update.xlsx", sheet = "PL Team Stats") %>% 
  mutate(Team = str_trim(Team))

sched <- sched %>%
  mutate(
    Team = str_trim(as.character(Team)),
    Opponent = str_trim(as.character(Opponent)),
    Home_Away = str_trim(as.character(Home_Away))
  )

ratings <- ratings %>%
  mutate(Team = str_trim(as.character(Team)))

sched_sim <- sched %>% 
  left_join(
    ratings %>% rename_with(~ paste0("team_", .), -Team),
    by = c("Team" = "Team")
  ) %>%
  left_join(
    ratings %>% rename_with(~ paste0("opp_", .), -Team),
    by = c("Opponent" = "Team")
  ) %>%
  mutate( 
    hca_team = case_when(Home_Away == "H" ~ home_court_adv,
                         Home_Away == "A" ~ -home_court_adv,
                         TRUE             ~ 0
  ),
  hca_opp = -hca_team
)

sim_scores <- function(sched_sim, n_sims = 10000,
                                 min_poss = 55, max_poss = 85) {

  n_games <- nrow(sched_sim)

  poss_mean <- (sched_sim$team_Possessions_Mean + sched_sim$opp_Possessions_Mean) / 2
  poss_sd   <- sqrt(sched_sim$team_Possessions_SD^2 +
                    sched_sim$opp_Possessions_SD^2) / 2

  team_eff_mean <- (sched_sim$team_Off_Rtg_Mean + sched_sim$opp_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_team
  team_eff_sd   <- sqrt(sched_sim$team_Off_Rtg_SD^2 +
                        sched_sim$opp_Def_Rtg_SD^2) / 2

  opp_eff_mean  <- (sched_sim$opp_Off_Rtg_Mean + sched_sim$team_Def_Rtg_Mean) / 2 +
                   sched_sim$hca_opp
  opp_eff_sd    <- sqrt(sched_sim$opp_Off_Rtg_SD^2 +
                        sched_sim$team_Def_Rtg_SD^2) / 2

  poss <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(poss_mean, each = n_sims),
          sd   = rep(poss_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  poss <- pmin(pmax(poss, min_poss), max_poss)

  team_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(team_eff_mean, each = n_sims),
          sd   = rep(team_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  opp_eff <- matrix( 
    rnorm(n_sims * n_games,
          mean = rep(opp_eff_mean, each = n_sims),
          sd   = rep(opp_eff_sd,   each = n_sims)),
    nrow = n_sims, ncol = n_games
  )

  team_score <- round(poss * team_eff / 100) 
  opp_score  <- round(poss * opp_eff  / 100)

  team_score[team_score < 0] <- 0
  opp_score[opp_score < 0]   <- 0

  win <- (team_score > opp_score) * 1L

  out <- tibble::tibble( 
    sim_id      = rep(seq_len(n_sims), times = n_games),
    game_index  = rep(seq_len(n_games), each  = n_sims),
    possessions = as.vector(poss),
    Team_Score  = as.vector(team_score),
    Opp_Score   = as.vector(opp_score),
    Win         = as.vector(win)
  )

  sched_rep <- sched_sim[rep(seq_len(n_games), each = n_sims),
                         c("Game_ID","Date","Team","Opponent","Home_Away")]

  dplyr::bind_cols(sched_rep, out %>% dplyr::select(-game_index))
}

sim_results <- sim_scores(sched_sim, n_sims=10000)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
library(scales)

record_dist <- sim_results %>% ## table showing the percentage of sims that had the team winning certain amounts of games
  group_by(sim_id) %>%
  summarise(wins = sum(Win), losses = n() - wins, .groups = "drop") %>%
  count(wins) %>%
  mutate(prob = n / sum(n)) %>%
  arrange(wins)

record_dist %>%
  arrange(wins) %>%
  mutate(
    Probability = percent(prob, accuracy = 0.1),
    'Sim Count' = n
  ) %>%
  select('Conference Wins' = wins, 'Sim Count', Probability) %>%
  kbl(
    format = "html",
    caption = "Projected Patriot League Record Distribution (10,000 Monte Carlo Simulations)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(3, bold = TRUE)
```

```{r, echo=FALSE}
matchup_summary <- sim_results %>% ## table showing the average score and win probability for each game
  group_by(Opponent, Home_Away) %>%
  summarise(
    avg_team_pts = mean(Team_Score),
    avg_opp_pts = mean(Opp_Score),
    avg_margin = mean(Team_Score - Opp_Score),
    win_prob = mean(Win),
    .groups = "drop"
  ) %>%
  arrange(desc(win_prob))

matchup_pretty <- matchup_summary %>% ## rounds everything to one decimal place for readability
  mutate(
    `Navy Pts` = round(avg_team_pts, 1),
    `Opponent Pts`  = round(avg_opp_pts, 1),
    `Avg Margin`    = round(avg_margin, 1),
    `Win %`         = round(win_prob * 100, 1)
  ) %>%
  select(
    Opponent, Site = Home_Away,
    `Navy Pts`, `Opponent Pts`, `Avg Margin`, `Win %`
  )

win_colors <- colorRampPalette(c("#f8d7da", "#fff3cd", "#d1e7dd"))(101) ## adds a red -> green color scale
margin_colors <- colorRampPalette(c("#f8d7da", "#ffffff", "#d1e7dd"))(101)

win_idx <- pmin(pmax(round(matchup_pretty$`Win %`) + 1, 1), 101) ## links win probability to the color scale
mrg_idx <- pmin(
  pmax(round((matchup_pretty$`Avg Margin` + 20) / 40 * 100) + 1, 1),
  101
)

matchup_pretty %>%
 kbl(
    format = "html",
    caption = "Projected Game Outcomes (Average Score, Margin, and Win Probability)",
    align = "c"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "responsive"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(5, background = margin_colors[mrg_idx]) %>%
  column_spec(6, background = win_colors[win_idx], bold = TRUE) %>%
  footnote(
    general = "Color scale: green = more favorable, red = less favorable."
  )
```

# Updated Final Standings

Here is how the Patriot League looks after $16$ games of conference play:

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(knitr)

conf_standings <- data.frame(
  Place = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
  Team = c("Navy", "Colgate", "American", "Lehigh", "Boston", "Loyola", "Lafayette", "Holy Cross", "Bucknell", "Army"),
  Conf_Record = c("15-1", "11-5", "9-7", "9-7", "8-8", "7-9", "6-10", "5-11", "5-11", "5-11"),
  Overall_Record = c("23-6", "17-12", "16-13", "13-16", "13-16", "11-18", "9-20", "10-19", "8-21", "11-18")
)

kable(conf_standings, caption = "2025-26 Patriot League (Sixteen Games In)", col.names = c("Rank", "Team", "Conference Record", "Overall Record"), align = "c")
```

And if every team finishes with their most likely number of wins, the 2025-26 Patriot League standings will look like this (based on the first $16$ games):

```{r warning=FALSE, message=FALSE, echo=FALSE}
pred_standings <- data.frame(
  Place = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
  Team = c("Navy", "Colgate", "American", "Lehigh", "Boston", "Loyola", "Lafayette", "Army", "Holy Cross", "Bucknell"),
  Conf_Record = c("17-1", "12-6", "11-7", "10-8", "9-9", "8-10", "7-11", "6-12", "6-12", "5-13"),
  Overall_Record = c("25-6", "18-13", "18-13", "14-17", "14-17", "12-19", "10-21", "12-19", "11-20", "8-23")
)

kable(pred_standings, caption = c("2025-26 Patriot League Standings (Based on First Sixteen Games)"), col.names = c("Rank", "Team", "Conference Record", "Overall Record"), align = "c")
```

Compared to the previous simulation's results:

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(knitr)

conf_standings <- data.frame(
  Place = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
  Team = c("Navy", "American", "Colgate", "Army", "Boston U", "Loyola MD", "Lehigh", "Holy Cross", "Lafayette", "Bucknell"),
  Conf_Record = c("12-6", "10-8", "10-8", "9-9", "9-9", "9-9", "8-10", "7-11", "7-11", "6-12"), 
  Overall_Record = c("20-11", "17-14", "16-15", "15-16", "14-17", "13-18", "12-19", "12-19", "10-21", "9-22")
)

kable(conf_standings, caption = "2025-26 Projected Patriot League Standings (Based on Non-Conference Schedule)", col.names = c("Rank", "Team", "Conference Record", "Overall Record"), align = "c")
```

Overperforming: American ($+1$ win), Colgate ($+2$ wins), Lehigh ($+2$ wins), Navy ($+5$ wins)

Underperforming: Army ($-3$ wins), Bucknell ($-1$ win), Holy Cross ($-1$ win), Loyola ($-1$ win) 

As Expected: Boston, Lafayette
